### 2017-10-13 paper/evaluation.tex
Pre-print version
Fixed overwritten latex rules

\section{Objects of Analysis}
\label{sec:subjects}

We used \github{}'s search API~\cite{githubsearch} to identify
projects that satisfy the following criteria: (1) the primary language
is Java\footnote{In case of projects in multiple languages, the
  \github{} API considers the predominant language as the primary
  language.}, (2) the project has at least 100 stars, (3) the latest
update was on or after January 1st, 2016, and (4) the \emph{readme}
file contains the string \emph{mvn}.  We focused on Java for its
popularity.  Although there is no clearcut limit on the number of
\github{} stars~\cite{github-stars} to define relevant projects, we
observed that one hundred stars was enough to eliminate trivial subjects. The third
criteria serves to skip projects without recent activity. The fourth
criteria is an approximation to find Maven projects.\Comment{ The
  rationale is that if the string \emph{mvn} exists in the
  \emph{readme} file, it may represent a Maven call (\eg, to compile
or to test the project).} We focused on Maven for its popularity on
Java projects.  Important to highlight that, as of now, the
\github{}'s search API can only reflect contents from repository
statistics (\eg, number of forks, main programming language); it does
not provide a feature to search for projects containing certain files
(\eg{}, \emph{pom.xml}) in the directory structure.
Figure~\ref{fig:subject-query} illustrates the query to the \github{}
API as an HTTP request.   The result set is sorted
in descending order of stars.

\begin{figure}[t!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    showstringspaces=false
}
\vspace{2mm}
\begin{lstlisting}
https://api.github.com/search/repositories?q=language:java
 +stars:>=100+pushed:>=2016+mvn%20in:readme&sort=stars
\end{lstlisting}
  \caption{\label{fig:subject-query} Query to the \github{} API for
  projects that (1) use Java, (2) contains at least 100
  stars, (3) has been updated on January 1st, 2016 (or later), (4) contains
  the string \emph{mvn} in the \emph{readme} file.}
  \vspace{-5ex}  
\end{figure}

We used the following methodology to select projects for
analysis. After obtaining the list of potential projects from GitHub, we filtered those
containing a \pomf{} file in the root directory.\Comment{  A Maven project may
contain several sub-modules with multiple \pomf{} files.}
Then, considering this set of Maven projects, we
executed the tests for \SubjectsReruns{} times to discard those projects with
issues
in the build file and non-deterministic results observed from sequential executions.
As of August 25th 2017, our search criteria returned a total of \SubjectsGithub{}
subjects.
From this set of projects,
\SubjectsGithubNotMaven{} projects were not Maven or did not have a
\pomf{} in the root directory, 
\SubjectsGithubNotTestable{} projects were not considered because of
environment incompatibility
(\eg, missing\Comment{ required web browser or database management system}~DBMS),
\SubjectsGithubFlaky{} projects were discarded because of
``flaky tests''~\cite{luo-etal-fse2014}. A ``flaky'' test is a test that passes or fails under
the same circumstances leading to non-deterministic results.
As some of our experiments consist of running tests on different
threads, we ignored these projects as it would be impractical
to identify whether a test failed due to a race condition or some
other source of flakiness.
From the remaining \SubjectsGithubConsistant{} projects with
deterministic results, we eliminated \SubjectsGithubTooManyFailures{}
projects with \SuiteFailingThreshold{} or more failing tests as to
reduce bias. For the
remaining projects with failing tests, we used the JUnit's
\CodeIn{@Ignore} annotation to ignore failing tests.
Our final set of subjects contains \numSubjs{} projects.
Figure~\ref{fig:subjects} summarizes our sample set.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.27\textwidth]{results/piechart-subjs.pdf}
  \caption{\label{fig:subjects}We fetched \SubjectsGithub{} popular projects
  hosted on \github{}. From this initial sample, we ignored
  \SubjectsGithubNotMaven{} projects without Maven support,
  \SubjectsGithubNotTestable{} with missing dependencies,
  \SubjectsGithubFlaky{} projects with flaky tests, and
  \SubjectsGithubTooManyFailures{} projects had at least
  \SuiteFailingThreshold{} of failing tests. We considered
  \numSubjs{} projects to conduct our study.}
  \vspace{-1ex}
\end{figure}

%\subsection{Setup and Replication}
\label{sec:setup}

@@

%% We are interested in understanding the prevalence of time-consuming
%% test suites and main sources of execution cost. We want to understand
%% how the execution cost is distributed on test cases within a test
%% suite and how developers approach test execution. Based on that, we
%% study parallelization of testing frameworks and build systems.  More
%% precisely, we investigate how prevalent test parallelization is, the
%% potential for improving execution cost, issues of flakiness that
%% hinders the use of parallelization, and how to address those issues.
%% More specifically, we pose the following research questions:

%% The first research question addresses the prevalence of long-running
%% test suites. We are interested to know if costly test suites are
%% common in open-source projects.  The second research question
%% addresses the relationship of test cases and the overall execution
%% cost: we are interested to investigate how the execution time is
%% distributed among test cases.  In the third research question, we
%% investigate if developers consider low-level parallelism features
%% available out-of-the-box to amortize test execution (see
%% Section~\ref{sec:modes}). In addition, we want to identify what
%% configurations are often used and why they are more popular (if any).
%% The fourth research question addresses the impact of low-level
%% parallelism on test execution from projects in our sample set. We want
%% to identify subjects that already use test parallelization and compare
%% their performance in contrast to sequential execution. In addition, we
%% are interested in evaluating the performance of sequential test suites
%% with different parallelization settings.
%% Finally, the fifth research question discusses the limitations and
%% insights to overcome the pitfalls of parallelization.

%% \Comment{
%%     \Fix{distribution of execution time per test case. For each subject
%%     identified in the first research question, we investigate how
%%     balanced is the cost of the test suite in contrast to the cost of
%%     test cases and if there are subjects where the time cost is mostly
%%     dominated by a small fraction of test cases.} \Fix{The third research
%%     question addresses the distribution of regression tests according
%%     to the use of computational resources.  We are interested in
%%     investigating if regression test suites are CPU intensive and if there
%%     are opportunities to improve performance. The RQ4 addresses}
%%     \Fix{...elaborate...}

%%     The rationale is that if the time cost of a regression test is equally
%%     distributed among test cases, the execution cost could be potentially
%%     improved by running tests in parallel (in contrast to the scenario
%%     where only one test case dominates most of the execution time).
%% }

@@

%%\newcommand{\RQFeasibilityTwo}{What is the distribution of CPU and IO bound
%%regression test suites from the sample set?}
%%
%%\newcommand{\RQAdoptionOne}{How uniformly distributed is the execution time
%%across test cases in costly projects?}
%%
%%\newcommand{\RQAdoptionTwo}{How often developers use the parallelism features
%%from build systems to improve runtime performance?}

@@

@@

\begin{figure}[ht]
  \centering
  \begin{subfigure}{0.182\textwidth}
    \centering
    \includegraphics[width=\textwidth]{results/barplot-timecost.pdf}
    \caption{\label{fig:rq1-barplot}}
  \end{subfigure}%
  \hfill
  \begin{subfigure}{0.25\textwidth}
    \centering
    \includegraphics[width=\textwidth]{results/boxplot-timecost.pdf}
    \caption{\label{fig:rq1-boxplot}}
  \end{subfigure}%
  \caption{(a)\Comment{ Testing time grouped by time cost ($t$): short run
    ($t<1m$), medium run ($1m \le t < 5m$), and long run ($5m \le{}
    t$);}~Number of projects in each cost group and
    (b)~Distribution of running times per cost group.}
  \vspace{-1ex}
\end{figure}

It is important to note that we under-estimated running times as we missed test modules not enabled for
execution in the root \emph{pom.xml}.\Comment{Some projects may omit long-running tests on their default
execution.} For instance, the project \CodeIn{apache.maven-surefire}
runs all unit tests in a few seconds.  According to our criteria, this
project is classified as \shortg{} but a closer look reveals
that only smoke tests are executed in this project by default.
In this project, integration and system tests, which take longer to run, are only accessible via
custom parameters, which we do not handle in our experimental setup.
We enabled such parameters for this specific project and observed that
testing time goes to nearly 30 minutes.  For simplicity, we considered
only the tests executed by default.
\Comment{
Although failing tests may introduce noise in our measurements, we
carefully eliminated subjects with more than \SuiteFailingThreshold{} of
test failures. 
}
From the \numSubjs{} testable projects, \numSubjsPass{} successfully
executed all tests and \numSubjsFail{} reported some test failure.
From these \numSubjsFail{} subjects, only 11 subjects
have more than 5\% of failing tests (7.3\% on average).
\begin{center}
\vspace{1ex}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQFeasibilityOne{}:}~\emph{We conclude that
    time-consuming test suites are relatively frequent in
    open-source projects.  We found that \percentMedLongRunning{} of
    the \numSubjs{} projects we analyzed (\ie{}, nearly 1 in every 4
    projects) take at least 1 minute to run and
    \percentLongRunning{} of them take at least 5 minutes to run.\Comment{
      (\ie, \numMedLong{} projects from \medg{} and \longg{}).}}
\end{minipage}
}
\vspace{1ex}
\end{center}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=.48\textwidth]{results/testcost-distribution.pdf}
  \caption{\label{fig:time-distributions}Distribution of test case time per project.}%
  \vspace{-3ex}
\end{figure}

Section~\ref{sec:rqA} showed that medium and long-running projects are
not uncommon, accounting to nearly \percentMedLongRunning{} of the
\numSubjs{} projects we analyzed.  Research question \numRQFeasibilityTwo{}
measures the distribution of test costs in test suites.\Comment{ as to estimate
potential of obtaining speedups with parallelization.}  In
the limit, if cost is dominated by a single test from a large test
suite, it is unlikely that parallelization will be beneficial as a
test method is the smallest working unit in test frameworks.
%% However, avoiding
%% frequent context switches is another factor to consider.  For example,
%% assuming there are at least two CPUs available for execution, cost can
%% be cut in half if two tests in a large test suite dominate execution
%% time and these tests are assigned to different CPUs.
%% It is therefore important to
%% speedup regressing testing in open-source projects.\Comment{not only
%%   to huge projects as those from Google~\cite{google-tap,google-ci}
%%   and Microsoft~\cite{prasad-shulte-ieee-microsoft-ci}.}
%% For the case
%% where cost is distributed more evenly across test cases, one expects
%% that speedups will be a function of the number of cores.
%% These contradictory forces, pushing number of tests and cost
%% of each test up and down, make prediction of effectiveness challenging.
Figure~\ref{fig:time-distributions} shows the time distribution of
individual test cases per project.
We observed that the average median times (see dashed horizontal red
lines) were small, namely 0.08s
for \medg{} projects and 0.16s for \longg{} projects, and the standard deviations associated with each distribution were
relatively low.\Comment{ Figure~\ref{fig:sd} shows the number of
  projects within specific ranges of $\sigma$ values.} High values of
$\sigma$ are indicative of CPU monopolization. We found only a small number
of those. The highest value of $\sigma$ occurred in
\CodeIn{uber\_chaperone}, a project from the \longg{} group.
This project contains only 26 tests, 17 of which take less than 0.5s
to run, one of which takes nearly 3s to run, two of which take nearly
11s to run, four of which takes on average 3m to run, and two of which
take $\sim$8m to run.
For this project, 98.4\% of the execution cost is dominated by 20\% of
the tests; without these two costly tests this project would have been
classified as short-running.
We did not find other projects with such extreme time monopolization
profile.
Project \CodeIn{facebookarchive\_linkbench} is also classified as
long-running and has the second highest value of $\sigma$.
For this project, however, cost is distributed more smoothly across 98
tests, of which 8 (8.1\%) take more than 1s to run with the rest of
the tests running faster.

@@

%% Figure~\ref{fig:time-distributions} shows that the average median
%% times are similar for \medg{} and \longg{}-running test suites.

\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQFeasibilityTwo{}:}~\emph{Overall, results indicate that
  projects with a very small number of tests monopolizing end-to-end
  execution time were rare. Time most often is distributed evenly
  across test cases.}
\end{minipage}
}
\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.13\textwidth}
    \centering
    \includegraphics[width=.9\textwidth]{results/boxplots-testcases.pdf}
    \caption{\label{fig:size-testsuites}}
  \end{subfigure}
  \hspace{3ex}
  \begin{subfigure}{0.28\textwidth}
    \centering
    \includegraphics[width=.9\textwidth]{results/scatter-testcost.pdf}
    \caption{\label{fig:scattercost}}
  \end{subfigure}
  \caption{\label{fig:time-versus-size}(a) Size of test suites; (b)
    Size versus running time of test suites.}%
\end{figure}
\end{center}

%% We are interested to know whether
%% most of the execution cost of a subject is dominated by a small subset
%% of test cases or if the cost is nearly equally distributed.

%% We also evaluated the dispersion of time distributions (one
%% distribution per project) to answer research question \numRQFeasibilityTwo{}.  To
%% measure dispersion \emph{across} projects we used Relative Standard
%% Deviation (RSD)~\cite{everitt-book-stats-2010}.  Note that, if we were
%% to analyze each project in isolation, the standard deviation of a
%% distribution ($\sigma$) would suffice to quantify how dispersed the
%% (time) distribution is.  However, in our case, we would like to be
%% able to compare and summarize dispersion across projects.  The RSD,
%% which is obtained dividing the standard deviation by the mean ($\mu$)
%% of a distribution, provides such normalization effect.  This metric
%% provides a lower bound (zero) but not an upper bound (somewhere close
%% to 1).  The smaller (larger) the value of RSD the more (less) uniform
%% the distribution is.  Consequently, the lower the value of RSD the
%% more parallelizable a test suite should be.

%% \begin{figure}[h!]
%%   \centering
%%   \includegraphics[width=0.5\textwidth]{R/testcost.pdf}  
%%   \caption{\label{fig:relativesd}Distribution of RSD ($\sigma/\mu$)
%%     across projects.}
%% \end{figure}

%% Figure~\ref{fig:relativesd} shows the distribution of RSD across
%% medium and long-running projects.  Results show that the distribution
%% is skewed to the right indicating that test costs are relatively well
%% distributed in most costly projects we analyzed \Fix{$\leftarrow$
%%   confirm}.

%% analyzed the execution time
%% for the \numMedLong{} projects from the \longg{} and \medg{} groups
%% (see Section~\ref{sec:rqA}).
%% For each subject we calculated the
%% relative standard deviation of the test cases: we collected the
%% elapsed time of each individual test, calculated the standard
%% deviation, and divided by the mean. \Jbc{I need to clarify the
%%   relationship "well/bad-balanced" regression test and relative
%%   standard deviation}

%% Results indicated that \Fix{...elaborate...}. \Jbc{We may identify
%% different groups of subjects}\Fix{TODO: collect data + compute the
%% statistic, create a scatter plot to identify groups of subjects}

%% Regression tests that are well distributed may benefit from
%% parallelism since more tests executes at the same time while the
%% opposite scenario may require a different approach. In the later
%% scenario, executing tests in parallel may have insignificant impact
%% since a small subset of test cases dominates the execution.}

@@

%% Commented as we don't follow this pattern on the other RQs
\Comment{
Adoption concerns with the popularity of  parallelization schemes and
(\numRQAdoptionTwo) how developers involved in costly projects, not
using parallelization, perceive this technology.
}
\begin{itemize}
    \item \numRQAdoptionOne. \textbf{\RQAdoptionOne{}}
\end{itemize}

@@

\vspace{1ex}
\subsubsection{Dynamic checking}
\label{sec:rqC-1}

To find dynamic evidence of parallelism, we ran the test suites from
our set of \numMedLong{} projects to output all key-value pairs of
Maven parameters.  To that end, we used the option~\CodeIn{-X} to
produce debug output and the option~\CodeIn{-DskipTests} to skip
execution of tests.  We skipped execution of tests as we observed from
sampling that only bootstrapping the Maven process suffices to infer
which parallel configuration modes it uses to run the
tests.  It is also important to point that we used the default
configurations specified in the project.  We inferred parallel
configurations by searching for certain configuration parameters in
log files. According to Maven's
documentation~\cite{maven-surefire-plugin}, a parallel configuration
depends either on (1) the parameter \CodeIn{parallel} to define the
parallelism mode within a JVM followed by the parameter
\CodeIn{threadCount} or (2) the parameter
\CodeIn{forkCount}\footnote{This parameter is named \CodeIn{forkMode}
  in old versions of Maven Surefire.} to define the number of forked
JVMs.  As such, we captured, for each project, all related key-value
pairs of Maven parameters and mapped those pairs to one of the
possible parallelization modes.  For instance, if a given project
contains a module with the parameter
\CodeIn{<forkCount>1C</forkCount>}, the possible classifications are
\ForkSeq{} or \ForkParMeth{}, depending on the presence and the value
of the parameter \CodeIn{parallel}.  If the parameter
\CodeIn{parallel} is set to \CodeIn{methods} the detected mode will be
\ForkParMeth{}.  Large projects may contain several test suites
distributed on different Maven modules potentially using different
configurations.  For those cases, we collected the Maven output from
each module discarding duplicates as to avoid inflating counts for
configuration modes that appear in several modules of the same
project. For instance, if a project contains two modules using the
same configuration, we counted only one occurrence.
Considering
our set of \numMedLong{} projects, we found that only
\numProjectsPar{} of those projects had parallelism enabled
by default, with only configurations \ParClassSeqMeth{},
\ParClassParMeth{}, and \ForkSeq{} being used. Configurations
\ParClassParMeth{} and \ForkSeq{} were the most popular among these
cases. Note that these results under-approximate real usage of
parallelism as we used default parameters in our scripts to spawn the
Maven process.  That decision could prevent execution of particular
test modules. Table~\ref{tab:freqmodes-dynamic} shows the
\numProjectsPar{} projects we identified where parallelism is enabled by default in Maven.
\begin{wraptable}{r}{0pt}%0.525\linewidth

  \footnotesize
  \centering
  \setlength{\tabcolsep}{2.5pt}
    \begin{tabular}{llcr}
        \toprule
        \multirow{2}{*}{\emph{Group}} & \multirow{2}{*}{\emph{Subject}} & \emph{\# of} & \multirow{2}{*}{\emph{Mode}}\\%
                                      &                                 & \emph{modules} &\\%
        \midrule%
        Medium & Californium & 2/20 & C2\\%
        Medium & Chaos & 1/1 & C2\\%
        Long & \Comment{apache} Flink & 66/74 & FC0\\%
        Long & \Comment{apache logging-}Log4J2 & 25/28 & FC0\\%
        Long & \Comment{javaslang }Javaslang & 3/3 & C3\\%
        Medium & Jcabi \Comment{jcabi-github} & 1/1 & C3\\%
        Long & \Comment{hazelcast hazelcast-}Jet & 6/7 & FC0\\%
        Long & \Comment{apache} Mahout & 8/9 & FC0\\%
        Long & \Comment{jankotek} MapDB & 1/1 & C3\\%
        Medium & \Comment{apache} OpenNLP & 4/4 & FC0\\%
        Medium & \Comment{yegor256} Rultor & 1/1 & C3\\%
        Medium & \Comment{yegor256} Takes & 1/1 & C3\\%
        Long & \Comment{vavr-io} Vavr & 3/3 & C3\\%
        \bottomrule%
    \end{tabular}
    \caption{Subjects with parallel test execution enabled by default.}
    \label{tab:freqmodes-dynamic}
%    \vspace{-1ex}
\end{wraptable}
Column ``\emph{Subject}'' indicates the name of the project, column
``\emph{\# of modules}'' indicates the fraction of modules containing
tests that use the configuration of parallelism mentioned in column
``\emph{Mode}''.
We note that, considering these projects, the modules that do not use
the configuration cited use the sequential configuration \Seq{}.
For example, three modules (=28-25) from Log4J2 use sequential
configuration. It came as a surprise the observation that
no project used distinct configurations in their modules.

@@

\vspace{1ex}
\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQAdoptionOne{}:}~\emph{Results indicate that test
  suite parallelization is underused.  Overall, only
  \percentParallel{} of costly projects (\numProjectsParTotal{} out of \numMedLong)
  use parallelism.}
\end{minipage}
}
\end{center}
\vspace{1ex}

@@

\begin{enumerate}
\item How long does it take for tests to run in your environment? Can
  you briefly define your setup?
\item Do you confirm that your regression test suite does *not* run in parallel?
\item\label{questionThree} Select a reason for not using parallelization:
  \begin{enumerate}[label=\alph*)]
  \item I did not know it was possible
  \item I was concerned with concurrency issues
  \item I use a continuous integration server
  \item Some other reason. Please elaborate.
  \end{enumerate}
\end{enumerate}
%% \begin{enumerate}
%% 	\item How long does it take for test to run in your
%%		environment?
%%	\item Can you briefly define your setup?
%%	\item Do you confirm that your project does not run in
%%		parallel?
%%	\item Select a reason for not using paralellization:
%%		\begin{enumarate}
%%			\item I did not know it was possible;
%%			\item I was concerned with concurrency issues;
%%			\item I use a continuous integration server;
%%			\item Some other reason.
%%		\end{enumerate}
%% \end{enumerate}
%% One of the goals of the first questions is to identify potential
%% discrepancies between our experimental environment and the environment
%% of developers.  Overall, we found that \Fix{...}

Considering question 1, we confirmed that execution time was
compatible with the results we reported in Section~\ref{sec:rqA}.
Furthermore, \emailsCI{} of the participants indicated the use of
Continuous Integration (CI) to run tests, with \emailsDistributed{} of
these participants reporting that test suites are modularized and
those modules are tested independently in CI servers through different
parameters.  Those participants explained that such practice helps to
reduce time to observe test failures, which is the goal of speeding up
regression testing.  A total of \emailsLocal{} participants answered
that they do run tests in their local machines.  Note, however, that
CI does not preclude low-level parallelization.  For example,
installations of open-source CI tools (\eg{}, Jenkins~\cite{jenkins})
in dedicated servers would benefit from running tests faster through
low-level test suite parallelization.
% \emailsNotDescribed{} developers did not described their environment.

Considering question 2, the answers we collected indicated, to our
surprise, that six of the \emailsProjectsAnswered{} projects execute
tests in parallel.  This mismatch is justified by cases where neither
of our checks (static or dynamic) could detect presence of
parallelism.  A closer look at these projects revealed that one of
them contained a \emph{pom.xml} file encapsulated in a jar file
(similar case as reported in Section~\ref{sec:rqC-2}), in one of the
projects the participant considered that distributed CI was a form of
parallelism, and in four projects the team preferred to implement
parallelization instead of using existing features from the testing
framework and the build system~---~in two projects the team
implemented concurrency control with custom JUnit test runners and in
two other projects the team implemented concurrency within test
methods.  Note that, considering these four extra cases (ignored two
distributed CI cases), the usage of parallelization increases from
\percentParallel{} to \percentParallelUpdated{}.  We do not consider
this change significant enough to modify our conclusion about
practical adoption of parallelization (\numRQAdoptionOne{}).
%% , one runs a manually created Thread to run some
%% tests, and the other runs in parallel by using Java 8 collection
%% streams, that allows the developers to iterate over a list in
%% parallel.
%% did not confirmed, however,
%% the developers confirmed the need of an extra parameter at the command
%% line to execute in parallel.

@@

%% In those projects, parallelization is active without
%% passing any extra parameters.  Section~\ref{sec:rqC-1} describes in
%% detail the methodology we used to find these subjects.
%and
%Section~\ref{seq:rq6-tradeoffs}, we verified that both
%% executions produce the same outcome to eliminate noise from failing
%% tests.  To compute the speedup, we divide the time obtained in the
%% sequential execution by the time obtained from the default execution.
%% For instance, if a project runs the tests sequentially in $10m$ and
%% the same execution runs in $5m$ with parallelization enabled (default
%% execution), the speedup is two.

@@

%%  We are
%% interested in understanding how efficiency (\ie, testing cost) and
%% flakiness (\ie, failing tests) are affected when we run a test suite
%% with different parallel configurations.

>>>>

\section{Setup and Replication}
\label{sec:setup}

@@

@@

@@

\begin{figure}[ht]%
  \centering
  \subfigure[\label{fig:rq1-barplot}]{
  \includegraphics[height=4cm,width=3cm]{results/barplot-timecost.pdf}
  }
  \qquad
  \subfigure[\label{fig:rq1-boxplot}]{
    \includegraphics[height=4cm,width=4.2cm]{results/boxplot-timecost.pdf}
  }%
  \caption{(a)\Comment{ Testing time grouped by time cost ($t$): short run
    ($t<1m$), medium run ($1m \le t < 5m$), and long run ($5m \le{}
    t$);}~Number of projects in each cost group and
    (b)~Distribution of running times per cost group.}
\vspace{-5mm}
\end{figure}

@@

It is important to note that we under-estimated running times as we missed test modules not enabled for
execution in the root \emph{pom.xml}.\Comment{Some projects may omit long-running tests on their default
execution.} For instance, the project \CodeIn{apache.maven-surefire}
runs all unit tests in a few seconds.  According to our criteria, this
project is classified as \shortg{} but a closer look reveals
that only smoke tests are executed in this project by default.
In this project, integration and system tests, which take longer to run, are only accessible via
custom parameters, which we do not handle in our experimental setup.
We enabled such parameters for this specific project and observed that
testing time goes to nearly 30 minutes.  For simplicity, we considered
only the tests executed by default.
From the \numSubjs{} testable projects, \numSubjsPass{} successfully
executed all tests and \numSubjsFail{} reported some test failure.
From these \numSubjsFail{} subjects, only 11 subjects
have more than 5\% of failing tests (7.3\% on average).
\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQFeasibilityOne{}:}~\emph{We conclude that
    time-consuming test suites are relatively frequent in
    open-source projects.  We found that \percentMedLongRunning{} of
    the \numSubjs{} projects we analyzed (\ie{}, nearly 1 in every 4
    projects) take at least 1 minute to run and
    \percentLongRunning{} of them take at least 5 minutes to run.\Comment{
      (\ie, \numMedLong{} projects from \medg{} and \longg{}).}}
\end{minipage}
}
\end{center}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=.48\textwidth]{results/testcost-distribution.pdf}
  \caption{\label{fig:time-distributions}Distribution of test case time per project.}%
  \vspace{-5mm}
\end{figure}

Section~\ref{sec:rqA} showed that medium and long-running projects are
not uncommon, accounting to nearly \percentMedLongRunning{} of the
\numSubjs{} projects we analyzed.  Research question \numRQFeasibilityTwo{}
measures the distribution of test costs in test suites.\Comment{ as to estimate
potential of obtaining speedups with parallelization.}  In
the limit, if cost is dominated by a single test from a large test
suite, it is unlikely that parallelization will be beneficial as a
test method is the smallest working unit in test frameworks.
Figure~\ref{fig:time-distributions} shows the time distribution of
individual test cases per project.
We observed that the average median times (see dashed horizontal red
lines) were small, namely 0.08s
for \medg{} projects and 0.16s for \longg{} projects, and the standard deviations associated with each distribution were
relatively low.\Comment{ Figure~\ref{fig:sd} shows the number of
  projects within specific ranges of $\sigma$ values.} High values of
$\sigma$ are indicative of CPU monopolization. We found only a small number
of those. The highest value of $\sigma$ occurred in
\CodeIn{uber\_chaperone}, a project from the \longg{} group.
This project contains only 26 tests, 17 of which take less than 0.5s
to run, one of which takes nearly 3s to run, two of which take nearly
11s to run, four of which takes on average 3m to run, and two of which
take $\sim$8m to run.
For this project, 98.4\% of the execution cost is dominated by 20\% of
the tests; without these two costly tests this project would have been
classified as short-running.
We did not find other projects with such extreme time monopolization
profile.
Project \CodeIn{facebookarchive\_linkbench} is also classified as
long-running and has the second highest value of $\sigma$.
For this project, however, cost is distributed more smoothly across 98
tests, of which 8 (8.1\%) take more than 1s to run with the rest of
the tests running faster.

\begin{figure}[t!]%
  \centering
  \subfigure[\label{fig:size-testsuites}]{
  \includegraphics[height=4.25cm, width=2.1cm]{results/boxplots-testcases.pdf}
  }
  \qquad
  \subfigure[\label{fig:scattercost}]{
    \includegraphics[height=4.25cm]{results/scatter-testcost.pdf}
  }%
  \caption{\label{fig:time-versus-size}(a) Size of test suites; (b)
    Size versus running time of test suites.}%
  \vspace{-5mm}
\end{figure}

@@

\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQFeasibilityTwo{}:}~\emph{Overall, results indicate that
  projects with a very small number of tests monopolizing end-to-end
  execution time were rare. Time most often is distributed evenly
  across test cases.}
\end{minipage}
}
\end{center}

@@

\begin{itemize}
    \item \numRQAdoptionOne. \textbf{\RQAdoptionOne{}}
\end{itemize}

@@

\subsubsection{Dynamic checking}
\label{sec:rqC-1}

To find dynamic evidence of parallelism, we ran the test suites from
our set of \numMedLong{} projects to output all key-value pairs of
Maven parameters.  To that end, we used the option~\CodeIn{-X} to
produce debug output and the option~\CodeIn{-DskipTests} to skip
execution of tests.  We skipped execution of tests as we observed from
sampling that only bootstrapping the Maven process suffices to infer
which parallel configuration modes it uses to run the
tests.  It is also important to point that we used the default
configurations specified in the project.  We inferred parallel
configurations by searching for certain configuration parameters in
log files. According to Maven's
documentation~\cite{maven-surefire-plugin}, a parallel configuration
depends either on (1) the parameter \CodeIn{parallel} to define the
parallelism mode within a JVM followed by the parameter
\CodeIn{threadCount} or (2) the parameter
\CodeIn{forkCount}\footnote{This parameter is named \CodeIn{forkMode}
  in old versions of Maven Surefire.} to define the number of forked
JVMs.  As such, we captured, for each project, all related key-value
pairs of Maven parameters and mapped those pairs to one of the
possible parallelization modes.  For instance, if a given project
contains a module with the parameter
\CodeIn{<forkCount>1C</forkCount>}, the possible classifications are
\ForkSeq{} or \ForkParMeth{}, depending on the presence and the value
of the parameter \CodeIn{parallel}.  If the parameter
\CodeIn{parallel} is set to \CodeIn{methods} the detected mode will be
\ForkParMeth{}.  Large projects may contain several test suites
distributed on different Maven modules potentially using different
configurations.  For those cases, we collected the Maven output from
each module discarding duplicates as to avoid inflating counts for
configuration modes that appear in several modules of the same
project. For instance, if a project contains two modules using the
same configuration, we counted only one occurrence.
Considering
our set of \numMedLong{} projects, we found that only
\numProjectsPar{} of those projects had parallelism enabled
by default, with only configurations \ParClassSeqMeth{},
\ParClassParMeth{}, and \ForkSeq{} being used. Configurations
\ParClassParMeth{} and \ForkSeq{} were the most popular among these
cases. Note that these results under-approximate real usage of
parallelism as we used default parameters in our scripts to spawn the
Maven process.  That decision could prevent execution of particular
test modules. Table~\ref{tab:freqmodes-dynamic} shows the
\numProjectsPar{} projects we identified where parallelism is enabled by default in Maven.

Column ``\emph{Subject}'' indicates the name of the project, column
\begin{wraptable}{r}{5cm}
\caption{Subjects with parallel test execution enabled by default.}
\label{tab:freqmodes-dynamic}
\footnotesize
\centering
\setlength{\tabcolsep}{2.5pt}
\begin{tabular}{llcr}
\toprule
\multirow{2}{*}{\emph{Group}} & \multirow{2}{*}{\emph{Subject}} & \emph{\# of} & \multirow{2}{*}{\emph{Mode}}\\%
   & & \emph{modules} &\\%
\midrule%
Medium & Californium & 2/20 & C2\\%
Medium & Chaos & 1/1 & C2\\%
Long & \Comment{apache} Flink & 66/74 & FC0\\%
Long & \Comment{apache logging-}Log4J2 & 25/28 & FC0\\%
Long & \Comment{javaslang }Javaslang & 3/3 & C3\\%
Medium & Jcabi \Comment{jcabi-github} & 1/1 & C3\\%
Long & \Comment{hazelcast hazelcast-}Jet & 6/7 & FC0\\%
Long & \Comment{apache} Mahout & 8/9 & FC0\\%
Long & \Comment{jankotek} MapDB & 1/1 & C3\\%
Medium & \Comment{apache} OpenNLP & 4/4 & FC0\\%
Medium & \Comment{yegor256} Rultor & 1/1 & C3\\%
Medium & \Comment{yegor256} Takes & 1/1 & C3\\%
Long & \Comment{vavr-io} Vavr & 3/3 & C3\\%
\bottomrule%
\end{tabular}
\end{wraptable}
``\emph{\# of modules}'' indicates the fraction of modules containing
tests that use the configuration of parallelism mentioned in column
``\emph{Mode}''.
We note that, considering these projects, the modules that do not use
the configuration cited use the sequential configuration \Seq{}.
For example, three modules (=28-25) from Log4J2 use sequential
configuration. It came as a surprise the observation that
no project used distinct configurations in their modules.

@@

\begin{center}
\fbox{
\begin{minipage}{8cm}
  \textit{Answering \numRQAdoptionOne{}:}~\emph{Results indicate that test
  suite parallelization is underused.  Overall, only
  \percentParallel{} of costly projects (\numProjectsParTotal{} out of \numMedLong)
  use parallelism.}
\end{minipage}
}
\end{center}

@@

\begin{enumerate}
\item How long does it take for tests to run in your environment? Can
  you briefly define your setup?
\item Do you confirm that your regression test suite does *not* run in parallel?
\item\label{questionThree} Select a reason for not using parallelization:
  \begin{enumerate}[label=\alph*)]
  \item I did not know it was possible
  \item I was concerned with concurrency issues
  \item I use a continuous integration server
  \item Some other reason. Please elaborate.
  \end{enumerate}
\end{enumerate}

Considering question 1, we confirmed that execution time was
compatible with the results we reported in Section~\ref{sec:rqA}.
Furthermore, \emailsCI{} of the participants indicated the use of
Continuous Integration (CI) to run tests, with \emailsDistributed{} of
these participants reporting that test suites are modularized and
those modules are tested independently in CI servers through different
parameters.  Those participants explained that such practice helps to
reduce time to observe test failures, which is the goal of speeding up
regression testing.  A total of \emailsLocal{} participants answered
that they do run tests in their local machines.  Note, however, that
CI does not preclude low-level parallelization.  For example,
installations of open-source CI tools (\eg{}, Jenkins~\cite{jenkins})
in dedicated servers would benefit from running tests faster through
low-level test suite parallelization.

Considering question 2, the answers we collected indicated, to our
surprise, that six of the \emailsProjectsAnswered{} projects execute
tests in parallel.  This mismatch is justified by cases where neither
of our checks (static or dynamic) could detect presence of
parallelism.  A closer look at these projects revealed that one of
them contained a \emph{pom.xml} file encapsulated in a jar file
(similar case as reported in Section~\ref{sec:rqC-2}), in one of the
projects the participant considered that distributed CI was a form of
parallelism, and in four projects the team preferred to implement
parallelization instead of using existing features from the testing
framework and the build system~---~in two projects the team
implemented concurrency control with custom JUnit test runners and in
two other projects the team implemented concurrency within test
methods.  Note that, considering these four extra cases (ignored two
distributed CI cases), the usage of parallelization increases from
\percentParallel{} to \percentParallelUpdated{}.  We do not consider
this change significant enough to modify our conclusion about
practical adoption of parallelization (\numRQAdoptionOne{}).

@@

@@
